Key Provisioning for the Internet of Things via Light

### 2015-10-13 paper/paper.tex
some more stuff

This paper proposes an alternative method, which is transmitting initial keying material via light.
It is mainly targeted on private users without technical background and should provide them with a simple though secure process for initializing.

As visible in Figure~\ref{fig:overview}, the user utilizes a smartphone application to enter, encode and emit keying material using the flash light of the phone.
The mote receives the message via a light sensor and reconstructs the keys.
After successful initialization and verification, the mote saves the keying material permanently.
On subsequent reboots, it is read from memory and immediately usable without any further action required.

@@

We implemented a prototype for an Android smartphone and a mote running the Contiki OS, \cite{dunkels04contiki}.
Contiki is an open source, state-of-the-art operating system for IoT devices\footnote{http://www.contiki-os.org/}.
We also tested some techniques to make the transmission error-prone.
Besides, we examined how fast the whole transmission could become while maintaining a high chance of correctness.

The remainder of the paper is structured as follows:
Section~\ref{sec:related_work} contains some background for security in the context of the~IoT and alternative approaches for key initialization.
In section~\ref{sec:communication_protocol}, we propose our protocol for the smartphone and the mote to communicate via light.
Details of our implementations are presented in section~\ref{sec:implementation}, both for the emitting~(sec.~\ref{sub:android_app}) and the receiving~(sec.~\ref{sub:mote}) parts.
In section~\ref{sec:results_and_discussion}, we discuss our findings regarding the protocol and its limits.
We conclude our work and give an outlook on what could be done next in section~\ref{sec:future_work}.

\section{Background and Related Work}
\label{sec:related_work}

In order to make the IoT reality, a specialized network stack is needed.
Motes have special constraints regarding energy and memory resources, e.g. batteries should work for multiple years and a motes program memory will be very small\footnote{e.g. 48kb program flash for telosB mote, \cite{telosb}}.

The \textit{IPv6 over Low-Power Wireless Personal Area Networks (6LoWPAN)} protocol stack as described in \cite{palattella2013standardized} is tailoring these needs.
The criteria for IoT network stacks are, according to them: low power usage, high reliability and Internet-suitability.
Therefore, tailored implementations of stack layers are presented, like \textit{PHY of IEEE 802.15.4}, which is a power saving physical/link layer.
In order to be Internet-enabled regarding the need for huge amounts of addresses, IPv6 is utilized.
However, in order to save energy, the headers of IPv6 packets are heavily compressed in 6LoWPAN.

There are some proposals for specialized security measures in this stack, e.g. AKES, \cite{krentz15akes}.
AKES makes use of session keys and a neighborhood-exploration-technique to gain a sufficient security level while being memory saving.
Security is even obtained after reboots of motes, but it relies on predistributed keys on each mote.

The problem of wireless and user-friendly key predistribution / key provisioning remains.
There are multiple approaches to solve this by securing the radio channel in some way, e.g.:

\cite{chen2011over} proposes key provisioning using eliptic curve cryptography (ECC).
However, while meeting high security standards, software implementations of ECC require a lot of memory and energy while hardware implementations increase the per-unit cost.
The problem of provisioning (public) keys still remains, too.

In order to secure radio communication, \cite{kuo2007message} utilized a Faraday cage in which a batch of motes can be programmed without interleaving of messages from an attacker outside of the cage.
Additionally, another device outside of the cave jams radio waves to prevent malicious communication.

>>>>

This paper proposes an alternative method, which is transmitting initial keying material via light.
Our method mainly targets private users without technical background.
As visible in Figure~\ref{fig:overview}, the user utilizes a smartphone application to enter, encode and emit keying material using the flash light of the phone.
The mote receives the message via a light sensor and reconstructs the keys.
After successful initialization and verification, the mote saves the keying material permanently.
On subsequent reboots, it is read from memory and immediately usable without any further action required.

@@

We implemented a prototype for an Android smartphone and a mote running the Contiki OS, \cite{dunkels04contiki}.
Contiki is an open source, state-of-the-art operating system for IoT devices\footnote{http://www.contiki-os.org/}.
We also tested techniques to make the transmission reliable.
Besides, we examined how fast the whole transmission could become while maintaining a high chance of correctness.

The remainder of the paper is structured as follows:
Section~\ref{sec:related_work} contains some background for security in the context of the~IoT and alternative approaches for key initialization.
In Section~\ref{sec:communication_protocol}, we propose our protocol for the smartphone and the mote to communicate via light.
Details of our implementations are presented in Section~\ref{sec:implementation}, both for the emitting~(Section~\ref{sub:android_app}) and the receiving~(Section~\ref{sub:mote}) parts.
In Section~\ref{sec:results_and_discussion}, we discuss our findings regarding the protocol and its limits.
In Section~\ref{sec:future_work}, we conclude our work and give suggestions for future research.

\section{Related Work}
\label{sec:related_work}

There are multiple approaches to key provisioning that secure the radio channel in some way, e.g.:

\cite{chen2011over} proposes key provisioning using eliptic curve cryptography (ECC).
However, while meeting high security standards, software implementations of ECC require a lot of memory and energy while hardware implementations increase the per-unit cost.
Moreover, they do not explain how public keys are provisioned.

In order to secure radio communication, \cite{kuo2007message} utilized a Faraday cage in which a batch of motes can be programmed without leaking messages to attackers outside of the cage.
Additionally, another device outside of the cage jams radio waves to prevent malicious communication.
